Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2023-02-16 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd
new ROC analyses done

## Diagnosis

### ROC

@@

### Histogram by group

@@

>>>>

## Full cohort

The full cohort analyses include all participants.

### Diagnosis

#### ROC

@@

#### Histogram by group

@@

### Genotype

Participants with unknown genotype risk were excluded from this analysis.

#### ROC

```{r}
# Re-format diagnosis column
t1_df$`Genotype Risk` <- factor(t1_df$geno_risk,
                     levels = c("Not CF","Low","High","Unknown"),
                     ordered = T)
t1_df$`Genotype Risk`[t1_df$`Genotype Risk`=="Unknown"]= NA
```

```{r}
roc_cf <- roc(response= t1_df$`Genotype Risk`,
              predictor = t1_df$`Na/Cl Ratio`, 
              levels = c("Low","High"),
              quiet = T)
coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)

roc_non_cf <- roc(response= t1_df$`Genotype Risk`,
              predictor = t1_df$`Na/Cl Ratio`, 
              levels = c("Not CF","Low"),
              quiet = T)
coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)

The blue curve shows the ROC for distinguishing between low vs. high risk, and the green is the ROC curve for distinguishing between non-CF vs. low risk. The optimal threshold for distinguishing between low vs. high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

#### Histogram by group

Vertical blue line indicates the threshold for low vs. high risk and green indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.

```{r warning=FALSE}
ggplot(t1_df[!is.na(t1_df$`Genotype Risk`), ], 
       aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)) +
  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
  theme_bw()
```

## Limited cohort

The limited cohort analyses only included participants with sweat chloride < 60.

### Diagnosis

#### ROC

```{r}
# Exclude those with Cl >= 60
t1_df = t1_df %>% filter(Cl < 60)
```

#### Histogram by group

### Genotype

Participants with unknown genotype risk were excluded from this analysis.

#### ROC

```{r}
roc_cf <- roc(response= t1_df$`Genotype Risk`,
              predictor = t1_df$`Na/Cl Ratio`, 
              levels = c("Low","High"),
              quiet = T)
coords_cf <- round(coords(roc_cf, "best", best.method = "youden"), 3)

roc_non_cf <- roc(response= t1_df$`Genotype Risk`,
              predictor = t1_df$`Na/Cl Ratio`, 
              levels = c("Not CF","Low"),
              quiet = T)
coords_non_cf <- round(coords(roc_non_cf, "best", best.method = "youden"), 3)

The blue curve shows the ROC for distinguishing between low vs. high risk, and the green is the ROC curve for distinguishing between non-CF vs. low risk. The optimal threshold for distinguishing between low vs. high risk was `r coords_cf$threshold` (sensitivity = `r coords_cf$sensitivity`, specificity = `r coords_cf$specificity`) based on the Youden index. The optimal threshold for distinguishing between non-CF vs. low risk was `r coords_non_cf$threshold` (sensitivity = `r coords_non_cf$sensitivity`, specificity = `r coords_non_cf$specificity`) based on the Youden index. The mean AUC for the two curves is `r round(mean(c(roc_cf$auc,roc_non_cf$auc)),3)`.

#### Histogram by group

Vertical blue line indicates the threshold for low vs. high risk and green indicates the threshold for non-CF vs. low risk. Na/Cl ratio was log-transformed for easier visualization.

```{r warning=FALSE}
ggplot(t1_df[!is.na(t1_df$`Genotype Risk`), ], 
       aes(log(`Na/Cl Ratio`), fill = `Genotype Risk`)) +
  geom_histogram(alpha = 0.3, position = "identity",bins = 50) +
  geom_vline(xintercept = log(coords_cf$threshold), color = "blue") +
  geom_vline(xintercept = log(coords_non_cf$threshold), color = "green") +
  ylab("Count") + xlab("ln(Na/Cl Ratio)") +
  theme_bw()
```
